The Shape Matching Element Method: Direct Animation of Curved Surface Models

### 2021-01-27 paper/sections/intro.tex

NURBS (Non-uniform Rational B Splines) are a popular higher-order modelling primitive which 
are used for computer-aided design (CAD), computational fabrication and computer animation. 
NURBS primitives were the first geometric representation used for physics-based animation, yet,
despite over three decades of research, animation of NURBS objects remains a challenge.

Isogeometric Analysis (IGA) is a physics simulation methodology that uses the control variables of the NURBS model
as the degrees-of-freedom (DOFS) of the simulation itself. 
Unfortunately IGA approaches for volumetric objects still require background volumetric structures, typically
regular grids that make satisfaction of boundary conditions difficult (which makes collision resolution difficult)
or more complicated cut-cell grids which introduce non-trivial root finding problems into the mix. 
Crucially, these simulation schemes typically assume models arise from engineering applications and meet tight geometric
criteria such as that the mesh is manufacturable. 
These inputs are much cleaner than those produced by a typical modeller.

@@

Our approach is an extension of the recently developed Virtual Element Method (VEM) for solving PDEs on domains tiled with arbitrary polygons.
We establish a connection between VEM and the well-known shape matching simulation algorithm  which enables us 
to derive equations of motion for an arbitrary NURBS model using Lagrangian Mechanics. 
Importantly, we show how to replace volumetric data structures for integration with ray casting approaches which enables
our meshless approach to isogeometric elastodynamic physically-based animation of volumetric structures.

>>>>

NURBS (Non-uniform Rational B Splines) are a popular higher-order modelling primitive which 
are used for computer-aided design (CAD), computational fabrication and computer animation. 
NURBS primitives were the first geometric representation used for physics-based animation, yet,
despite over three decades of research, animation of curved models such as those built with NURBS, remains a challenge.

Isogeometric Analysis (IGA) is a physics simulation methodology that uses the control variables of the NURBS model
as the degrees-of-freedom (DOFS) of the simulation itself. 
Unfortunately IGA approaches for volumetric objects still require background volumetric structures, typically
regular grids that make satisfaction of boundary conditions difficult (which makes collision resolution hard)
or more complicated cut-cell grids which introduce non-trivial root finding problems into the mix. 
Crucially, these simulation schemes typically assume models arise from engineering applications and meet tight geometric
criteria such as that the mesh is manufacturable. 
These inputs are much cleaner than those produced by a typical animation modeller.

@@

Our approach draws inspiration from the recently developed Virtual Element Method (VEM) for solving partial differential equations on domains tiled with arbitrary polygons.
We establish a connection between VEM and the well-known shape matching simulation algorithm. This allows us 
to derive equations of motion for an arbitrary curved surface model, made of NURBS, via Lagrangian Mechanics. 
Importantly, we show how to replace volumetric data structures for integration and blending weight computation with ray casting approaches which enables
our meshless approach to elastodynamic physically-based animation.

### 2021-01-27 paper/sections/methods.tex

 \begin{equation}
    \defX\left(\refX\right) = \sum_{\prt_i\in\Prts} \bldxi\Pmatxi\Pi_i\vc{q}_i, 
    \label{eq:kinematics}
 \end{equation} where $\bldxi$, $\Pmatxi$, $\Pi_i$ and $\vc{q}_i$ are the blending weights, polynomial basis, projection operator and control points for the $i^{th}$
 part respectively.

 \subsection*{Multiple Part Projection}
As in ~\refsec{shapematching} we construct each $\Pi_i$ via shape matching. 
It is tempting to use an identical polynomial basis matrix, $\Pmatx$ for all parts, $\prt_i$, but recall that constructing $\Pmatx$ requires the selection of a deformation
origin $\bar{\refX}$~(\refeq{projection}). 
Using the same origin for all polynomials limits the expressivity of the kinematic model~\cite{STBS:2011}.
Instead we choose a number of points in $\refX$ ($|\Prts| > |\XBs| > 0$, where $\XBs$ is the set of points)  to act as deformation origins~(\refsec{origins}).
For each part we now construct $\Pmatxi$ using the associated deformation origin $\bar{\refX}_j\in \XBs$, where each origin can
be shared between multiple parts.

\begin{figure}[h]
    \includegraphics[width=\columnwidth]{figures/projection_operator_solve}
    \caption{A simple, four part model with two deformation origins and the corresponding matrix equation for the quadratic coefficients.}
    \label{fig:multiparts}
\end{figure}

>>>>

 \begin{equation}
    \defX\left(\refX\right) = \sum_{\prt_i\in\Prts} \bldxi\Pmatxi\Pi_i\vc{q}_i, 
    \label{eq:kinematics}
 \end{equation} where $\bldxi$, $\Pmatxi$, $\Pi_i$ and $\vc{q}_i$ are the blending weights, polynomial basis, projection operator and control points for the $i^{th}$
 part respectively.  
 \ty{I'm wondering if we should add a line emphasizing that while we linearly blend, the resulting deformation map is a nonlinear map unique to each X. But maybe this obvious}

 \subsection*{Multiple Part Projection}
As in ~\refsec{shapematching} we construct each $\Pi_i$ via shape matching. 
It is tempting to use an identical polynomial basis matrix, $\Pmatx$ for all parts, $\prt_i$, but recall that constructing $\Pmatx$ requires the selection of a deformation
origin $\bar{\refX}$~(\refeq{projection}). 
Using the same origin for all polynomials limits the expressivity of the kinematic model~\cite{STBS:2011}.
Instead we choose a set of points $\XBs$ ($0 < |\XBs| < |\Prts|$) to act as deformation origins~(\refsec{origins}).
For each part we now construct $\Pmatxi$ using its associated deformation origin $\bar{\refX}_j\in \XBs$, where each origin can
be shared between multiple parts.

\begin{figure}[h]
    \includegraphics[width=\columnwidth]{figures/projection_operator_solve}
    \caption{A simple, four part model with two deformation origins and the corresponding matrix equation for the quadratic coefficients. \ty{these $T^1$ and $T^2$ blocks in the U matrix need to be swapped I think} }
    \label{fig:multiparts}
\end{figure}

### 2021-01-27 paper/sections/related.tex

%% Set the scene
Geometric modelling is a necessary precursor to physics-based animation, however differing geometric representations for modelling and simulation often requires
time-consuming, complex geometry processing pipelines. 
For instance, the popular tetrahedral finite element approach for simulating solid elastodynamics~\cite{10.1145/2343483.2343501} requires robust algorithms for converting input surface geometry into 
a volumetric tetrahedral mesh. This is a difficult problem and while significant progress has been made,
even the most robust volumetric methods~\cite{Hu:2018:TMW:3197517.3201353} can be time consuming, fail to maintain correspondence between the input model and output simulation mesh,
and don't work directly on curved surface representations such as NURBS.

@@

\emph{Isogeometric Analysis}~\cite{10.1002/9780470749081.ch7} endeavors to perform simulation directly on the NURBS output from 
Computer-Aided Design (CAD) software. Initial attempts used NURBS surfaces to represent the mid-surface of thin objects~\cite{10.1145/176579.176580}.
Volumetric simulations relied on volumetric NURBS~\cite{10.1007/978-3-642-03596-8_2} but were limited to a narrow class of geometries.
Finite volume methods are applicable to more general geometries~\cite{HEINRICH20121645,10.1002/nme.2311} but require a volumetric mesh to be generated.
Modern approaches are constructed around the extended finite element method which enriches the standard finite element basis with discontinuous basis functions
to enable improved boundary handling~\cite{https://doi.org/10.1002/nme.3120,10.1007/s00466-013-0854-7,https://doi.org/10.1002/nme.4852,SAFDARI2016373,10.1145/3355089.3356576}. 
These methods typically start with an easy-to-generate structured volumetric mesh (tetrahedral or hexahedral), ``cutting'' the NURBS geomtric model against
it to enable boundary handling (such a mesh is called a \emph{cutcell} mesh). 
Like volumetric meshing, this cutting operation can be difficult and our ideal method would avoid it. 
Some cutcell algorithms assume engineering/manufacturing quality input, which puts tight requirements on input models~\cite{10.1145/3355089.3356576}.
Finally these approaches require additional mechanisms to ensure that simulation results lie inside the shape space of the input model's primitives,
which increases the complexity.

@@

To alleviate these restrictions we turn to other meshless methods popular in computer graphics and 
engineering~\cite{10.1145/1028523.1028542,https://doi.org/10.1002/fld.1650200824,10.1145/1778765.1778776,10.1145/2010324.1964968, 10.1145/1944846.1944855}.
These methods support more advanced constitutive models and integration schemes but often require background integration meshes and lose the direct connection with
modelling geometry. 
Like cutcell-based, Isogeometric Analysis approaches, additional constraints must be added to ensure that simulation output can be represented by the input model.
Given this, we conclude that there is no existing algorithm that meets our desiderata for success.

Our algorithm takes the Shape Matching approach as inspiration, but rather than follow the PBD formalism, we interpret Shape Matching as a 
Virtual Element Method (VEM)~\cite{10.1142/S0218202512500492,10.1142/S021820251440003X}. Virtual Elements are an extension of 
mimetic finite differences~\cite{10.1142/S0218202505000832,10.1016/j.jcp.2013.07.031} to weak-form variational problems. 
VEM relaxes the mesh generation requirements of the finite element method by enabling the solution of partial differential equations
on domains partitioned with arbitrary polytopes~\cite{10.1145/3386569.3392389}. 
The solution inside each polytope is approximated using a polynomial function of a specified order.
VEM typically assumes that the boundary of the the problem domain is described by a piecewise linear complex
which makes its standard formulation incompatible with our NURBS based input geometry.

\subsection*{Contributions} 
In this paper we develop a new Shape Matching-based, Virtual Element Method which is directly compatible 
with NURBS input geometry rather than piecewise linear surfaces. 
Furthermore, we improve the expressivity of the VEM basis by using a shape blending approach inspired by
algorithms for skinning~\cite{skinningcourse:2014}.
Our method is entirely meshless (requiring no volumetric meshes or cutcells) and is compatible with standard constitutive models and time integrators.
It guarantees that simulation output is directly consumable by the input modelling software and can ingest models which include large gaps,
intersections and disconnected primitives without additional user input. 
These features mean that our algorithm is the first truly meshless approach for the direct simulation of NURBS models for physics-based animation.

>>>>

%% Set the scene
Geometric modeling is a necessary precursor to physics-based animation, however connecting differing geometric representations for modelling and simulation often requires
time-consuming, complex geometry processing pipelines. 
For instance, the popular tetrahedral finite element approach for simulating solid elastodynamics~\cite{10.1145/2343483.2343501} requires robust algorithms for converting input surface geometry into 
a volumetric tetrahedral mesh. This is a difficult problem and while significant progress has been made,
even the most robust volumetric methods~\cite{Hu:2018:TMW:3197517.3201353} can be time consuming, fail to maintain correspondence between the input model and output simulation mesh,
and don't work directly on curved surface representations such as NURBS.

@@

\emph{Isogeometric Analysis}~\cite{10.1002/9780470749081.ch7} endeavors to perform simulation directly on the NURBS output from 
Computer-Aided Design (CAD) software. Initial attempts used NURBS surfaces to represent the mid-surface of thin objects~\cite{10.1145/176579.176580}.
Volumetric simulations relied on volumetric NURBS~\cite{10.1007/978-3-642-03596-8_2} but were limited to a narrow class of geometries.
Finite volume methods are applicable to more general geometries~\cite{HEINRICH20121645,10.1002/nme.2311} but require a volumetric mesh to be generated.
Modern approaches are constructed around the extended finite element method which enriches the standard finite element basis with discontinuous basis functions
to enable improved boundary handling~\cite{https://doi.org/10.1002/nme.3120,10.1007/s00466-013-0854-7,https://doi.org/10.1002/nme.4852,SAFDARI2016373,10.1145/3355089.3356576}. 
These methods typically start with an easy-to-generate structured volumetric mesh (tetrahedral or hexahedral), ``cutting'' the NURBS geometric model against
it to enable boundary handling (such a mesh is called a \emph{cutcell} mesh). 
Like volumetric meshing, this cutting operation can be difficult and our ideal method would avoid it. 
Some cutcell algorithms assume engineering/manufacturing quality input, which puts tight requirements on input models~\cite{10.1145/3355089.3356576}.
Finally these approaches require additional mechanisms to ensure that simulation results lie inside the shape space of the input model's primitives,
which increases the complexity.

@@

To alleviate these restrictions we turn to other meshless methods popular in computer graphics and 
engineering~\cite{10.1145/1028523.1028542,https://doi.org/10.1002/fld.1650200824,10.1145/1778765.1778776,10.1145/2010324.1964968, 10.1145/1944846.1944855}.
These methods support more advanced constitutive models and integration schemes but often require background integration meshes, limit themselves to low order deformation functions 
and lose the direct connection with
modelling geometry. 
Like cutcell-based, Isogeometric Analysis approaches, additional constraints must be added to ensure that simulation output can be represented by the input model.
Given this, we conclude that there is no existing algorithm that meets our desiderata for success.

Our algorithm takes the Shape Matching approach as inspiration, but rather than follow the PBD formalism, we interpret Shape Matching as a 
Virtual Element Method (VEM)~\cite{10.1142/S0218202512500492,10.1142/S021820251440003X}. Virtual Elements are an extension of 
mimetic finite differences~\cite{10.1142/S0218202505000832,10.1016/j.jcp.2013.07.031} to weak-form variational problems. 
VEM relaxes the mesh generation requirements of the finite element method by enabling the solution of partial differential equations
on domains partitioned with arbitrary polytopes~\cite{10.1145/3386569.3392389}. 
The solution inside each polytope is approximated using a polynomial function of a specified order.
VEM typically assumes that the boundary of the the problem domain is described by a piecewise linear complex
which makes its standard formulation incompatible with our curved input geometry.

\subsection*{Contributions} 
In this paper we develop a new Shape Matching-based, Virtual Element Method which is directly compatible 
with NURBS input geometry rather than piecewise linear surfaces. 
Furthermore, we improve the expressivity of the VEM basis by using a shape blending approach inspired by
algorithms for skinning~\cite{skinningcourse:2014}.
Our method is entirely meshless (requiring no volumetric meshes or cutcells) and is compatible with standard constitutive models and time integrators.
It guarantees that simulation output is directly consumable by the input modelling software and can ingest models which include large gaps,
intersections and disconnected primitives without additional user input. 
These features mean that our algorithm is the first truly meshless approach for the direct simulation of curved surface, NURBS models for physics-based animation.

### 2021-01-27 paper/sections/results.tex

\dave{we should show the nurbs model in rhino/fusion for every example, maybe an exploded view as well}

\dave{mention that not so many standard NURBS models so we modelled lots of these ourselves .. show's power of approach that we can model and sim all these examples.}

\dave{use as many different meshes as possible for figures (i.e try not to have the same mesh show up in more than one figure if possible).}

\dave{we should report which models we made ourselves}

\begin{figure*}
  \includegraphics[width=\textwidth]{figures/chicken.pdf}
  \caption{Simulation of a chicken model with gaps between surfaces.}
  \label{fig:chicken}
\end{figure*}

\reftbl{perf} shows the size of all our examples along with performance statistics. 
Note that the individual parts of all models are \emph{not connected}, and no continuity constraints or constructive solid geometry operations have been applied. 
Rather, the SEM approach implicitly couples the boundary parts together to allow for seamless volumetric simulation. 
Models were created using Autodesk Fusion 360 \cite{AutodeskFusion360} and Rhinocerous 3D 7 \cite{mcneel-rhinoceros}.

For raycasting operations and rendering, we triangulate the NURBS surfaces, which only represents an implementation simplification. Our algorithm is not dependent on triangulating the boundary, and an ideal alternative would be using a path tracer.

\ty{Would be nice to mention TNURBS, which would give us a good reason to include the wrench (which i really wanna do). I doubt we have time to do a 3 point bending test, but a simple elastic wrench simulation would be easy to cook up}
For handling trimmed NURBS surfaces, we augment our sampling of the surfaces by using raycasting integration in 2D. Trimmed NURBS may be defined by a boundary curve in the parametric space, so we raycast to find intersection intervals in the UV space on which quadrature points are generated. For rendering Trimmed NURBS, we discretize the the boundary curve as a polyline and use triangle \cite{shewchuk96b, SHEWCHUK200221} to form a boundary conforming triangulation. The vertices of this triangulation serve as fixed set of high resolution UV samples for constructing a render mesh.

\begin{table*}[h]
  \caption{Performance of the Shape Matching Element Method on all Examples. \dave{need table that also reports material params/constitutive model or add it here}}
  \label{tbl:perf}
  \begin{center}
  \begin{tabular}{l c c c c c c c c}
   \textbf{Example} & \textbf{CPU} & $|\vc{q}|$ &  \textbf{Sample Model} & \textbf{Quadrature} & \textbf{Weights} & \textbf{Build $\Pi$} & \textbf{Time Step} \\
   \hline 
   \rowcolor[HTML]{DAE8FC} 
   \textbf{Cantilever}      & ? & ? & ? & ? & ? & ? & ? & ? \\
   \textbf{Coffee mug}      & ? & ? & ? & ? & ? & ? & ? & ? \\
   \textbf{Rocket}          & ? & 648 & ? & ? & ? & ? & ? & ? \\
   \textbf{Castle}          & ? & ? & ? & ? & ? & ? & ? & ? \\
   \textbf{Chicken}         & ? & ? & ? & ? & ? & ? & ? & ? \\
   \textbf{Starship}        & ? & ? & ? & ? & ? & ? & ? & ? \\
   \textbf{Tire}            & ? & ? & ? & ? & ? & ? & ? & ? \\
   \hline
  \end{tabular}
  \end{center}

@@

@@

%% Results that relate to simulation output
%correctness
We first validate the physical plausibility of our method using a small 2D patch test (see \reffig{patchtest}). 
Here the test object is a square made of four edges (not joined at the corners) simulated using linear VEM with a single deformation center.
We apply a battery of boundary conditions and resolve the deformation of the element by minimizing the elastic potential.
We note that SEM is able to represent rigid motions, as well as shearing and anisotropic stretching. 
This implies that, with sparse weights, SEM can resolve these motions locally, leading to physically plausible simulation results.

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/patch_test.pdf}
  \caption{2D patch test. By applying affine transformations to the boundary of the original undeformed model (left), we show that solving the static problem gives rigid motion for rotation and constant strain for shearing and stretching (right).}

  % \caption{2D patch test. Left: the original undeformed model defined by four surface elements on the boundary. Right:  applying affine transformations to the boundary and show that solving the static problem gives rigid motion for rotation and constant strain for shearing and stretching}
  \label{fig:patchtest}
\end{figure}

We further demonstrate the convergence of our SEM with respect to linear tetrahedral finite elements when increasing the number of patches in use.
In \reffig{convergence}, we compare identical cantilevered beams, simulated using a Neohookean constitutive model (Young's Modulus $=$ $0.001$ GPa, Poisson's Ratio $=$ 0.45) and implicit time integration with a timestep of $0.01$s. 
We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of 24 independent NURBS patches, shows good agreement with FEM. Each subdivision of the NURBS beam enables more complicated kinematics, but very few surface elements are needed to produce compelling results.

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/beams.pdf}
  \caption{Our SEM simulation is able to converges to high resolution FEM simulation result (left) as we increase the number of surfaces in the model (right). }
  % \caption{(left) high res FEM cantilevered bar (right) several simulations using more and more surface.}
  \label{fig:convergence}
\end{figure}

We also show that our raycasting weight computation is able to create shape-aware output. \reffig{independence} shows that manipulating
parts that are nearby but separated will behave in an appropriately independent fashion. 
\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/independent_movement}
  \caption{Our raycasting weight computation produces share-aware blending weights. Here the locality of our blending weights allows the two legs of the Gumby model to move independently. }
  \label{fig:independence}
\end{figure}

% relaxed modelling constraints
By virtue of its meshless nature, our SEM is robust to a wide range of challenging models with large gaps, intersections and disconnected primitives.
\reffig{badmodels} shows simulations of two jelly coffee mugs. 
The top row shows a model with negligible overlap, whereas the bottom row shows the result of a careless modeller who has deeply embedded the handle of the mug in the body in order to attach it, creating a large area of overlap.
In both cases SEM produces a plausible physically-based animation without requiring additional model clean-up.
\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/mug_overlap}
  \caption{Our SEM is robust to overlapping regions and self-intersections. Here the simulation result of the coffee mug with overlapping handle (bottom) still matches its corresponding counterpart without negligible overlaps (top). }
  % \caption{Coffee mug with overlapping handle. (top) row of models with increasing overlap (middle) single weight image showing behaviour in the overlapping region, (bottom) simulation result for each one}
  \label{fig:badmodels}
\end{figure}
Similarly, in \reffig{chicken} we show a simulation with gaps between surfaces. Despite the lack of explicit connectivity, the blending weights have the effect of implicitly enforcing connections at these seams.

%different material parameters / material models 
Leveraging the fact that the equations of motion for our SEM are derived using a general elastic potential, our SEM is theoretically guaranteed to support arbitrary constitutive models.
% We derive the equations of motion for SEM using a general elastic potential and thus SEM supports arbitrary constitutive models. 
\reffig{rocket} shows a rocket ship simulated with both high stiffness and low stiffness materials. 
%multiple materials
Oue SEM also supports the simulation of objects made up of heterogenous materials (see \reffig{tire}).
\begin{figure*}[htp]
  \includegraphics[width=\textwidth]{figures/rocket.pdf}
  \caption{Our SEM is able to handle models with open surfaces. Here we show a rocket (with open surfaces at the bottom) lands on the ground under gravity. }
  \label{fig:rocket}
\end{figure*}

@@

% different time integration scheme
\begin{figure}[h]
  \includegraphics[width=\columnwidth]{example-image-a}
  \caption{show mug simulation using newton's method, linearly implicit Euler \Honglin{Maybe I can do that quickly}. }
  \label{fig:time_integration}
\end{figure}

%editing example
For downstream applications, the output of our SEM simulation is itself editable in a NURBS modelling program (see \reffig{edit}). 
\begin{figure}[h]
  \includegraphics[width=\columnwidth]{example-image-a}
  \caption{show input to simulation, show output then show output loaded in rhino/fusion 360 }
  \label{fig:edit}
\end{figure}

%show stoppers
Finally we show the ability of SEM to handle complex simulations involving deformation, contact and friction.
In \reffig{staypuft} and \reffig{starship}, we show two staypuft characters collide with each other in space and a Space-V starship lands on Mars in a graceful way.
\begin{figure*}[htp]
  \includegraphics[width=\textwidth,height=3in]{example-image-a}
  \caption{sequence of frames from staypuft simulation}
  \label{fig:staypuft}
\end{figure*}

>>>>

\begin{table*}[h]
  \caption{Performance of, and parameters for the Shape Matching Element Method on all examples. All timings are reported in seconds, physical  parameters are
  reported with appropriate units. $\rho$ is the applied density, \textbf{E} is the  Young's Modulus and $\mu$ is the Poisson's ratio.
  }
  \label{tbl:perf}
  \begin{center}
  \begin{tabular}{l c c c c c c c c c c}
   \textbf{Example} $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(GPa) & $\mu$ &\textbf{Sample Model}(s) & \textbf{Quadrature}(s)& \textbf{Weights}(s)& \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
   \hline 
   \rowcolor[HTML]{DAE8FC} 
   \textbf{Cantilever}      & ? & ? & ? & ? & ? & ? & ? & ? & ? & ? \\
   \textbf{Coffee mug}      & ?& ? & ? & ? & ? & ? & ? & ? & ? & ?  \\
   \textbf{Rocket}          & ?& ? & 648 & ? & ? & ? & ? & ? & ? & ?\\
   \textbf{Castle}          & ?& ? & ? & ? & ? & ? & ? & ? & ? & ?\\
   \textbf{Chicken}        & ? & ? & ? & ? & ? & ? & ? & ? & ? & ?\\
   \textbf{Starship}       & ? & ? & ? & ? & ? & ? & ? & ? & ? & ?\\
   \textbf{Tire}           & ? & ? & ? & ? & ? & ? & ? & ? & ? & ?\\
   \hline
  \end{tabular}
  \end{center}

@@

%\dave{we should show the nurbs model in rhino/fusion for every example, maybe an exploded view as well}

%\dave{mention that not so many standard NURBS models so we modelled lots of these ourselves .. show's power of approach that we can model and sim all these examples.}

%\dave{use as many different meshes as possible for figures (i.e try not to have the same mesh show up in more than one figure if possible).}

%\dave{we should report which models we made ourselves}
We implemented SEM in a combination of MATLAB and C++ using both GPToolbox~\cite{gptoolbox} and Bartels~\cite{bartels} for geometry processing and constitutive models 
respectively. Benchmarks were performed using a  ~\dave{CPU} with \dave{xx}MB of RAM.

\reftbl{perf} shows the size of all our examples along with performance statistics and relevant parameters.
Note that the individual parts of all models are \emph{not connected}, and no continuity constraints or constructive solid geometry operations have been applied. 
Rather, the SEM approach implicitly couples the boundary parts together to allow for seamless volumetric simulation. 
Models were created using Autodesk Fusion 360 \cite{AutodeskFusion360} and Rhinocerous 3D 7 \cite{mcneel-rhinoceros}.

For raycasting operations and rendering we rely on Embree~\cite{10.1145/2601097.2601199} and triangulate the NURBS surfaces, which only represents an implementation simplification. 
Our algorithm is not dependent on triangulating the boundary, and an ideal alternative would be using a path tracer.

\nicetohave{Would be nice to mention TNURBS, which would give us a good reason to include the wrench (which i really wanna do). I doubt we have time to do a 3 point bending test, but a simple elastic wrench simulation would be easy to cook up}
For handling trimmed NURBS surfaces, we augment our sampling of the surfaces by using raycasting integration in 2D. Trimmed NURBS may be defined by a boundary curve in the parametric space, so we raycast to find intersection intervals in the UV space on which quadrature points are generated. For rendering Trimmed NURBS, we discretize the the boundary curve as a polyline and use triangle \cite{shewchuk96b, SHEWCHUK200221} to form a boundary conforming triangulation. The vertices of this triangulation serve as fixed set of high resolution UV samples for constructing a render mesh.

@@

%% Results that relate to simulation output
%correctness
\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/patch_test.pdf}
  \caption{2D patch test. By applying affine transformations to the boundary of the original undeformed model (left), we show that solving the static problem gives rigid motion for rotation and constant strain for shearing and stretching (right).}
  % \caption{2D patch test. Left: the original undeformed model defined by four surface elements on the boundary. Right:  applying affine transformations to the boundary and show that solving the static problem gives rigid motion for rotation and constant strain for shearing and stretching}
  \label{fig:patchtest}
\end{figure}

We first validate the physical plausibility of our method using a small 2D patch test~(\reffig{patchtest}). 
Here the test object is a square made of four edges (not joined at the corners) simulated using linear polynomials with a single deformation center.
We apply a battery of boundary conditions and resolve the deformation of the element by minimizing the elastic potential.
We note that SEM is able to represent rigid motions, as well as shearing and anisotropic stretching. 
This implies that, with sparse weights, SEM can resolve these motions locally, leading to physically plausible simulation results.

We further demonstrate \emph{qualitative} convergence of SEM with respect to linear tetrahedral finite elements when increasing the number of patches in use.
In \reffig{convergence}, we compare identical cantilevered beams, simulated using a Neohookean constitutive model (Young's Modulus $=$ $0.001$ GPa, Poisson's Ratio $=$ $0.45$) and implicit time integration with a timestep of $0.01$s. 
We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of 24 independent NURBS patches, shows good agreement with FEM. Each subdivision of the NURBS beam enables more complicated kinematics, 
but very few surface elements are needed to produce compelling results.

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/beams.pdf}
  \caption{Our SEM simulation is able to qualitatively converge to a high resolution FEM simulation result (left) as we increase the number of surfaces in the model (right). }
  \label{fig:convergence}
\end{figure}

We also show that our raycasting weight computation is able to create shape-aware output. \reffig{independence} shows that manipulating
parts that are nearby but separated will behave in an appropriately independent fashion. Our grumpy model is capable of extending 
its leg without causing unrealistic deformations in the plant foot.

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{figures/independent_movement}
  \caption{Our raycasting weight computation produces share-aware blending weights. Here the locality of our blending weights allows the two legs of the grumpy model to move independently. }
  \label{fig:independence}
\end{figure}

% relaxed modelling constraints
By virtue of its meshless nature, SEM is robust to a wide range of challenging models with large gaps and disconnected primitives.
\reffig{fig:chicken} shows frames from a simulation of a rubber chicken. Note that the chicken model itself features large gaps between the individual NURBS parts. 
Despite the lack of explicit connectivity, the SEM blending weights have the effect of implicitly enforcing connectivity at these seams.

\begin{figure*}[htp]
  \includegraphics[width=\textwidth]{figures/chicken.pdf}
  \caption{Simulation of a chicken model with gaps between surfaces.}
  \label{fig:chicken}
\end{figure*}

SEM is also robust to intersections in modelling input. \reffig{badmodels} shows simulations of two jelly coffee mugs. 
The top row shows a model with negligible overlap, whereas the bottom row shows the result of a careless modeller who has deeply embedded the handle of the mug in the body in order to attach it, creating a large area of overlap.
In both cases SEM produces a plausible physically-based animation without requiring additional model clean-up.

\begin{figure}[H]
  \includegraphics[width=\columnwidth]{figures/mug_overlap}
  \caption{Our SEM is robust to overlapping regions and self-intersections. Here the simulation result of the coffee mug with overlapping handle (bottom) still matches its corresponding counterpart without negligible overlaps (top). }
  % \caption{Coffee mug with overlapping handle. (top) row of models with increasing overlap (middle) single weight image showing behaviour in the overlapping region, (bottom) simulation result for each one}
  \label{fig:badmodels}
\end{figure}

%different material parameters / material models 
Since the equations of motion for SEM are derived using a general elastic potential, it is theoretically capable of supporting arbitrary elastic constitutive models.
% We derive the equations of motion for SEM using a general elastic potential and thus SEM supports arbitrary constitutive models. 
\reffig{rocket} shows a rocket ship simulated with both high stiffness (steel) and low stiffness materials (rubber). 
In both cases intuitive and visually pleasing results are created wherein the trajectories correctly reflect the desired material characteristics.  
%multiple materials
\begin{figure*}[htp]
  \includegraphics[width=\textwidth]{figures/rocket.pdf}
  \caption{SEM is able to produce animation using a wide variety of material parameters. Here we simulate both a steel and rubber rocket ship. 
  To make this even more challenging, the rocket model is an open surface (at the bottom), but a plausible animation is still generated.}
  \label{fig:rocket}
\end{figure*}

SEM also supports the simulation of objects made up of heterogenous materials. By specifying different material properties inside nested 
parts we are able to simulate a drag racing wheel with a metal hub and soft rubber tread~(\reffig{tire}).

@@

% different time integration scheme
\nicetohave{Integration Figure}
%\begin{figure}[h]
%  \includegraphics[width=\columnwidth]{example-image-a}
%  \caption{show mug simulation using newton's method, linearly implicit Euler \Honglin{Maybe I can do that quickly}. }
%  \label{fig:time_integration}
%\end{figure}

%editing example
For downstream applications, the output of  SEM simulation is itself editable in a NURBS modelling program (see \reffig{edit}). 
This can be useful for visual effects pipelines, allowing artists to easily post process animations using the same tools in which they were created.
\begin{figure}[h]
  \includegraphics[width=\columnwidth]{example-image-a}
  \caption{Here we show the output of an SEM animation loaded into Rhinoceros 3D 7 for additional editing.}
  \label{fig:edit}
\end{figure}

%show stoppers
Finally we show a few additional examples, highlighting the ability of SEM to handle complex simulations involving deformation, contact and friction.
In \reffig{staypuft} we show two astronauts in jaunty space suits collide in zero gravity while in \reffig{starship} 
we show an animation of the Space-V starship landing on Mars in a graceful way.

\begin{figure*}[htp]
  \includegraphics[width=\textwidth,height=3in]{example-image-a}
  \caption{Two astronauts, in jaunty spacesuits collide while spacewalking in zero-g.}
  \label{fig:staypuft}
\end{figure*}
